Machine Learning Autoencoder Applied to Communication Channels

### 2019-06-19 Rapport/Article/Article.tex
Seance 06/19/2019

For a classification problem in communication, the \textit{categorical cross-entropy} loss function $l(\textbf{p},\textbf{q}):\mathbb{R} ^{N_L}\times \mathbb{R} ^{N_L}\mapsto \mathbb{R}$ defined in \eqref{eq:c-e} is most common. It is derived from the log-likelihood of a training set where $q_j$ is the estimated probability of the outcome $j$ and $p_{j}$ is the true probability. Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{murphy2013machine} where $j=1,...,N_L$.
\begin{equation}\label{eq:c-e}
	l(\textbf{p},\textbf{q})=-\sum _{j}p_{j}\log \left( q_{j}\right)
\end{equation}

@@

The architecture of the DNN necessary to achieve a great performance in terms of BER is available in table \ref{tab:arraydecoder}. What should be remarked is the first layer of the DNN. A \textit{Lambda} layer is a layer in which the activation operations are user defined and the parameters are non-trainable. Thus, we were able to define a layer that act as a BSC to force our DNN to learn its effects in the training data. The fact that the parameters are non-trainable means that the noise layer is applied in forward propagation only.

The choice of the crossover probability of the training channel, $p_{t}$, was based in the work of \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the best in terms of decoding precision after the DNN is trained. In addition, a deeper NN showed best results in \cite{DBLP:journals/corr/OSheaH17} and were adopted to solve the problem. The training parameters are displayed in table \ref{tab:arraydecoderTrain}.

Trainin set for array decoding.

The choices of training parameters is a mix between what other works in literature uses and a hyper parametric analytic approach, where several trials lead to the decision. First, the binary cross-entropy as loss function is widely used in literature and  showed to converge to the desired result consistently during tests. For the optimizer, the \textit{Adam} built-in Keras optimizer was chosen. Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}. Next, a batch size matching the size of the training set which is sufficiently small to still operate in the regime of small batch was chosen. In \cite{DBLP:journals/corr/KeskarMNST16}, the authors demonstrated better performance for a NN training with stochastic gradient descent and its variants operating in a small-batch regime, usually between 32-512 data points. Finally, the necessary number of epochs to training is sensitive to the problem, since it impacts directly the training computational cost. As a result, we tried several training epoch sizes and found $2^{16}$ to have precise results and be computationally viable.

@@

@@

Trainin set for one-hot decoding, say how was pseudo-randomly generated, sa that the stastistic was checked to have 1/2 of 1s and 1/2 of 0s.

@@

However, we still need a decoding matrix.... Introduces autoencoder that could train without G

@@

@@

\begin{table}
\caption{DNN array autodecoder training parameters.} 
\begin{center}
\begin{tabular}{ l  |  l | l | l  }
    \thickhline
 Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
 Binary cross-entropy & Adam & $2^{14}$ & $256$ \\
    \thickhline
  \end{tabular}
    \label{tab:arrayautoencoderTrain}
\end{center}
\end{table}

@@

Explain any different training procedure

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
    \caption{One hot decoding BER performance. DNN one-hot decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
\end{figure}

\subsection{Autoencoder}
BER Curves of simulation varying p, best choice of p, analyse autoencoder with best choice of p

@@

compare time in decoding between map, autoencoder and a decoder.

>>>>

For a classification problem in communication, the \textit{categorical cross-entropy} loss function $l(\textbf{p},\textbf{q}):\mathbb{R} ^{N_L}\times \mathbb{R} ^{N_L}\mapsto \mathbb{R}$ defined in equation \eqref{eq:c-e} is most common. It is derived from the log-likelihood of a training set where $q_j$ is the estimated probability of the outcome $j$ and $p_{j}$ is the true probability. Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{murphy2013machine} where $j=1,...,N_L$. Another commonly used loss function, indispensable for this research, is the \textit{mean squared error} (MSE). The MSE is the  sum of squared distances between our target variable and predicted values and it is defined in equation \ref{eq:mse}.

\begin{align}
\
	\text{Cross-entropy: }l_1(\textbf{p},\textbf{q}) &=-\sum _{j}p_{j}\log \left( q_{j}\right)\label{eq:c-e} \\
	\text{MSE: } l_2(\textbf{p},\textbf{q}) &=\frac{1}{N_L}\sum _{j}(p_{j}-q_{j})^{2} \label{eq:mse}
\end{align}

@@

The architecture of the DNN necessary to achieve a great performance in terms of BER is not evident. What should be remarked though is the first layer of the DNN. A \textit{Lambda} layer is a layer in which the activation operations are user defined and the parameters are non-trainable. Thus, we were able to define a layer that act as a BSC to force our DNN to learn its effects in the training data. The fact that the parameters are non-trainable means that the noise layer is applied in forward propagation only.

The choice of the crossover probability of the training channel, $p_{t}$, was based in the work of \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the best in terms of decoding precision after the DNN is trained. In addition, a deeper DNN showed best results in \cite{DBLP:journals/corr/OSheaH17} where 3 hidden layers were used. Consequently, the same architecture were adopted to solve the problem as shown in table \ref{tab:arraydecoder}.

@@

The training set for the array decoder contained all the possible messages as training labels and its correspondents code words as training data. Because of the chosen encoding, the training data is a $256\times16$ matrix and the training label is a $256\times8$ matrix. To generate a possible message, a stochastic algorithm was developed to fabricate random 8-length vectors of integers 1s and 0s which the average value considering all elements was around $0.5$. In this way we could guarantee that our training environment were representative of a real life application. Where different bits are transmitted with the same probability.

Even though it is not recommended to train a NN with all the available data - because it might lead to over-fitting -, for this case does not seems to be a problem. WHY????? (noisy channel? onehot decodeR?)

The choices of training parameters is a mix between what other works in literature uses and a hyper parametric analytic approach, where several trials lead to the decision. First, the binary cross-entropy as loss function is widely used in literature and  showed to converge to the desired result consistently during tests. For the optimizer, the \textit{Adam} built-in Keras optimizer was chosen. Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}. Next, a batch size matching the size of the training set which is sufficiently small to still operate in the regime of small batch was chosen. In \cite{DBLP:journals/corr/KeskarMNST16}, the authors demonstrated better performance for a NN training with stochastic gradient descent and its variants operating in a small-batch regime, usually between 32-512 data points. Finally, the necessary number of epochs to training is sensitive to the problem, since it impacts directly the training computational cost. As a result, we tried several training epoch sizes and found $2^{16}$ to have precise results and be computationally viable. The training parameters are displayed in table \ref{tab:arraydecoderTrain}.

@@

The training set in this case differs from the array decoder. The training labels are one-hot vectors which maps into real messages through a bijective function. Thus, it is a matrix of dimension $256\times256$ composed of 256 different on-hot vectors. The training data is still the linear encoding of the real messages by the matrix $\textbf{G}$.

@@

Despite of this, an encoding matrix is still needed. A commonly used method in ML for image recognition without training labels is the autoencoder. In this context, the NN is able to learn not only a decoding algorithm, but also a decoding one as introduced in the following section.

@@

\subsubsection{Array Autoencoder}

@@

\begin{table}
\caption{DNN array autodecoder training parameters.} 
\begin{center}
\begin{tabular}{ l  |  l | l | l  }
    \thickhline
 Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
 MSE & Adam & $2^{17}$ & $256$ \\
    \thickhline
  \end{tabular}
    \label{tab:arrayautoencoderTrain}
\end{center}
\end{table}

\subsubsection{One-hot Autoencoder}

Briefly state that it is basically the same, apart from the input and output layers AND it demands a tremendously different architecture.

\begin{table}
\caption{DNN one-hot autoencoder architecture.} 
\begin{center}

  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: 256 \\
 & Dense: X, activation: ReLU  \\
 & Dense: 16, activation: Sigmoid  \\ \hline

 \multirow{2}{*}{Decoder} & Dense: 128, input size: 16 \\
 & Dense: 256, activation: Sigmoid  \\ \hline

    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
    \thickhline
  \end{tabular}
    \label{tab:onehotautoencoder}
\end{center}
\end{table}

\begin{table}
\caption{DNN one-hot autodecoder training parameters.} 
\begin{center}
\begin{tabular}{ l  |  l | l | l  }
    \thickhline
 Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
 MSE & Adam & $2^{17}$ & $256$ \\
    \thickhline
  \end{tabular}
    \label{tab:onehotautoencoderTrain}
\end{center}
\end{table}

@@

This section is divided in three parts. The first shows the results obtained to the two implemented decoders. Observations during the training of the DNN are made. Secondly, the results for the autoencoders is brought with some observations during the training period. The discussion and analysis for these two subsections is around the BER performance for the decoders and autoencoders. Finally, a decoding time performance analysis is explained and the results are shown and discussed.

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
    \caption{One hot decoding BER performance. DNN one-hot decoder trained without a noisy channel.}\label{fig:1HD}
\end{figure}

\subsection{Autoencoders}
BER Curves of simulation varying p, best choice of p, analyse autoencoder with best choice of p

@@

say that only the decoding part was measured  for autoencoders
